---
title: "My first paper"
site: bookdown::bookdown_site

author: |
  | Anna Quaglieri  ^[Corresponding author: anna.quaglieri16@gmail.com] $^{1,2}$, Soroor Zadeh $^{1,2}$
  | $^1$Walter and Eliza Hall Institute,  $^2$University of Melbourne 

abstract: |
  Hopefully with this minimal example you will appreciate the robustness and flexibility of `bookdown` and you will start writing amazing reproducible documents!!

bibliography: biblio.bib
link-citations: yes

output: 
  bookdown::pdf_document2:
    toc: yes
    keep_tex: true

rmd_files: ["01-main.Rmd", "02-Additional_files.Rmd", "03-References.Rmd"]

fontsize: 12pt
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  
  \usepackage{setspace}\doublespacing
  \usepackage{lineno}
  \linenumbers
---

\newpage

```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.height = 5,fig.width = 5,fig.path = "img/",tidy = TRUE, cache = TRUE)

library(bookdown)
library(knitr)
library(tidyverse)
library(png)
library(grid)
library(kableExtra)
```


# Introduction
\label{sec:intro}

This is my introduction to the flexible `bookdown` [@bookdown] package.

`bookdown` extends the syntax provided by R Markdown, allowing automatic numbering of figures / tables / equations, and cross-references within a single documents.

# Results
\label{sec:res}

All the Figures created by the chunks in this article will be saved into the `img/` which is created in the setup chunk at the beginning of this document. 

```r
knitr::opts_chunk$set(echo = TRUE,fig.height = 5,fig.width = 5,fig.path = "img/",tidy = TRUE)
```


## Generate figures within a chunk 

```{r bell-curve, fig.cap="Red trees in autumn",eval=TRUE,fig.width=3,fig.height=3}
plot(density(rnorm(100)))
```

It's a pretty cool bell curve in Figure \ref{fig:bell-curve} \@ref(fig:bell-curve). 

* Don't use `_` in labels! This is taken from `bookdown` book: *If you want to cross-reference figures or tables generated from a code chunk, please make sure the chunk label only contains alphanumeric characters (a-z, A-Z, 0-9), slashes (/), or dashes (-).*

## Insert `png` figures

```{r loud-kooka, fig.cap="Loud Kookabara",eval=TRUE,fig.height=4,fig.width=4}
img <- png::readPNG("img/kooka.png")
grid.raster(img)
```


## Insert `pdf` figures with `knitr::include_graphics`

```{r bookdown-docs, fig.cap="All types of documents with `bookdown`",eval=TRUE,fig.height=4,fig.width=6}
knitr::include_graphics("img/reproducible-documents.pdf")
```


## Insert tables with `knitr::kable`

```{r tree-diamonds,eval=TRUE,results='hold'}
tab <- knitr::kable(head(diamonds),
      caption = 'A table of the first 10 rows and three columns of the diamonds data.',
      booktabs = TRUE)

kableExtra::kable_styling(tab, latex_options = 'hold_position')
```

We created the simple Table \@ref(tab:tree-diamonds). See more on tables and long tables https://bookdown.org/yihui/bookdown/tables.html. 


## Extra tips: Setup your bibliography with `utils::citation`

```{r extra,eval=TRUE,echo=TRUE}
print(utils::citation("rticles"),bibtex=TRUE)
```

## Reference/Re-number files in supplementary material

Have a look at the beautiful autumn trees in Figure \@ref(fig:red-trees).

# Conclusions

Hopefully now you have a pretty good overview of the options offered by `bookdown`! Also, the best way to learn is to go through other people's GitHub accounts and have a look at how they do things as well as ask questions!

\newpage






