---
title: "title"
site: bookdown::bookdown_site

author: |
  | Author 1^[Corresponding author: email@email.com] $^1$, Author 2 $^1$, Author 3 $^2$
  | $^1$Affiliation1,  $^2$Affiliation2

abstract: |
  Your abstract goes here...

bibliography: biblio.bib

output: 
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
fontsize: 12pt
header-includes: 
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{setspace}\doublespacing
  \usepackage{lineno}
  \linenumbers
---


\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.height = 5,fig.width = 5,fig.path = "img/")

library(bookdown)
library(knitr)
library(tidyverse)
```


# Introduction
\label{sec:intro}

This is my intro. All of this uses the amazing [@bookdown] package.

Bookdown extends the syntax provided by R Markdown, allowing automatic numbering of figures / tables / equations, and cross-referencing them.

https://stackoverflow.com/questions/25824795/how-to-combine-two-rmarkdown-rmd-files-into-a-single-output

http://landscapeportal.org/blog/2017/09/06/r-markdown-template-for-a-scientific-manuscript/


# Results
\label{sec:res}

Amazing results after introducing the story in Section \ref{sec:intro}.

All the Figures created by the chunks in this article will be saved into the `img/` which is created in the setup chunk at the beginning of this document. 

## About the Normal Distribution
\label{sec:res-trees}

* Problem of “floating environments” in LaTeX

```{r bell-curve, fig.cap="Red trees in autumn",eval=TRUE}
plot(density(rnorm(100)))
```

It's a pretty cool bell curve in Figure \ref{fig:bell-curve}. Whatch out for `_` in cross-referencing labels!


If you want to cross-reference figures or tables generated from a code chunk, please make sure the chunk label only contains alphanumeric characters (a-z, A-Z, 0-9), slashes (/), or dashes (-).

It is always good to have some Greek letters in your paper! Like Equation \@ref(eq:binom) below:

\begin{equation} 
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}
  (\#eq:binom)
\end{equation} 


## About the scatter plot

```{r scatter, fig.cap="Loud Kookabara",eval=TRUE}
plot(1:10,1:10)
```


```{r tree-diamonds,eval=TRUE}
kable(head(diamonds),
      caption = 'A table of the first 10 rows and three columns of the diamonds data.',
      booktabs = TRUE)
```

We created the simple Table \ref{tab:tree-diamonds}. See more on tables and long tables https://bookdown.org/yihui/bookdown/tables.html. 

You can also see a scatter plot in Figure \@ref(fig:sup1).


## Extra tips: Setup your bibliography

```{r extra,eval=FALSE}
print(citation("rticles"),bibtex=TRUE)
```


# Conclusions

Bla Bla....


\newpage


